\vspace{-2cm} Government websites as data: \\ A methodological pipeline with application to the websites of municipalities in the United States

### 2018-09-01 paper/manuscript.tex
spelling and grammar edits

Local governments convey voluminous information about all aspects of their policymaking, policy implementation, and public deliberation, via their official websites. The vital role of official websites in connecting the government and the governed has motivated a wave of research on the contents of government websites \citep[e.g.,][]{grimmelikhuijsen2010transparency,wang2005evaluating,osman2014cobra}. Despite the potential for automated scraping of website contents, the conventional approach to data collection in projects focused on government websites involves manual content extraction from each website in the dataset. Though highly accurate, the manual approach to data collection is costly, and cannot be scaled to capture even a fraction of the volume of content available on government websites. In this paper we present a methodological pipeline that can be used to automatically scrape government websites in order to build datasets that can be used for text analysis. We provide an illustrative application in which we explore the ways in which the textual contents on city government websites in six American states (IN, LA, NY, WA, CA and TX) correlate with the partisanship of the city mayor.

Though there exists a variety of software tools that are designed to automatically scrape all of the files available at a website \citep{glez2013web}, raw website downloads have to be processed significantly before the files are adequately prepared for content (e.g., text, image) analysis. We describe and provide solutions to two central challenges in automatically gathering and analyzing website textual contents. First, plain text must be extracted from the files. This involves purging the files of syntax in HTML or other markup and programming languages, and discarding any character encoding errors that result from reading in files with incorrectly specified character encodings. This challenge would arise in any context in which researchers sought to study the textual contents of websites, and is not unique to comparative analysis of government websites. The second challenge we address in our methodological pipeline is, however, specific to the research objective of comparing websites on the basis of a common lexicon. For any two governments, the textual signatures that most dramatically differentiate the textual contents of their websites consist of what we can call ``boilerplate'' text---header, footer, or other titling text that is designed to identify the website as being associated with a specific government entity (e.g., ``Welcome to the city of Santa Cruz'', ``The City of Los Angeles welcomes you''). This boilerplate text is replicated across many files that are associated with a government's website, but it provides little information regarding the form and/or function of the government. The second methodological innovation we offer in our pipeline is designed to minimize the impact of this boilerplate text on the comparative analysis of government website content.

Government websites provide information about how public policies shape the lives of local residents, and how local residents can engage with government to shape public policy. As such, government websites reflect both the results of, and inputs to, the political leadership in the city. In our illustrative application we explore the ways in which the contents of city government websites differ on the basis of the partisanship of the city's elected executive. A substantial body of research has found that the partisanship of the mayor affects city governance along multiple dimensions, including city budget priorities \citep{de2016mayoral}, policies affecting inequality in cities \citep{einstein2016mayors}, and framing of criminal justice policy \citep{marion2013mayor}. Furthermore, recent media coverage of changes to government websites that follow transitions in party control suggest that changes in web content are salient government actions, as perceived by the general public \citep{sharfstein2017science,kirby2017trump,duarte2017deniable}. We study whether significant differences between city governments based on mayoral partisanship are reflected in the contents of city websites.

@@

 According to \cite{Mayhew1974}, politicians engage in advertising, credit claiming and position taking in order to get re-elected. Official city websites allow mayors to perform all three of these functions. Their offices frequently take a prominent position on the front page, and many websites also feature a picture of the mayor. We present an example of this in Figure \ref{fig:eriemayor}. The Erie, Pennsylvania website homepage presents an image of Democratic mayor, Joseph Schember, along with a list of laudable attributes of the city. In local politics, where campaign funds are low, this lends incumbents a crucial advantage in becoming more well-known among their constituencies. Furthermore, municipal politics gives incumbents clear and tangible achievements they can point to, such as completed infrastructure projects, the acquisition of federal or state funding, or the hosting of city-wide events. City websites present an opportunity for local officials to brandish these accomplishments. Finally, they also give mayors a platform from which they can advertise their political beliefs. On municipal websites, this may not manifest in the form of brazen partisanship, but more subtle avenues are available. As noted by \cite{einstein2016mayors}, there are stark differences in the spending preferences of Democratic and Republican mayors. City websites can then be used to communicate the stance of a mayor on social or economic programs. Another advantage of websites with regard to communication is that unlike direct social interactions, officials have full control over them.

@@

The existing research that uses scraped websites provides an indication of the theoretical value of empirical analysis of web contents produced by governments, public officials, and candidates for office. The most pertinent literature to our research is the e-governance literature, which focuses on the online presence of governments from a usability and public service point of view. For the most part, research in this category develops a classification scheme to rate websites in terms of accessibility, ease-of-use and function, and then hand-codes a set of websites according to these criteria \citep[e.g., ][]{Urban2002,Armstrong2011,Feeney2017}. As an example, \cite{grimmelikhuijsen2012developing} study local government websites with the goal of uncovering how they aid the goal of transparency. To this end, they analyze a set of Dutch municipalities in which air quality had deteriorated. The authors test whether local governments provide citizens with information about potential complications and solutions associated with this issue. Like most e-government studies however, this publication does not make any use of automated text analysis.
%Websites have also played a major role in the field of media studies, as scholars have scraped and analyzed the online presence of newspapers, as well as the more diffuse world of online political blogs \citep{Adamic2005,Gentzkow2010}. \cite{Lin2011} provide a good example for a study which makes extensive use of automated content analysis - a necessity arising from its dataset of 66830 blog posts and 57221 online news articles. The authors estimate the political slant of these entities by counting the frequencies with which politicians of either side are mentioned and determine that blogs are generally more biased. Unfortunately for us, the authors don't go into the details of their text analysis, and offer no information on the acquisition and pre-processing of the data.
%Another well-known example fitting into this area of study is the set of studies conducted by King et al. \citep{KING2013,King2014,KING2017}, in which the authors study censorship by the country's government on its lively blogosphere. However, the authors also provide no information on how their data was collected ``our extensive engineering effort, which we do not detail here for obvious reasons [...]''.

The websites of politicians and their parties have also fallen under scholarly scrutiny. Researchers have found that in order to identify the constituencies, motives and modes of communication of these actors, their websites can be very illuminating sources of information \citep{Druckman2009,Druckman2010,Cryer2017,Esterling2011,Esterling2011a,Norris2003,Therriault2010}. \cite{Druckman2009,Druckman2010} rely on the \textit{National Journal} to find the websites, then hand-coded them. \cite{Cryer2017} mentions that she relied on Archive-it, a webservice of the Internet Archive. Though the Internet Archive may provide extensive coverage of high profile officials or national governments, we found that its coverage of municipal government websites was sparse and irregular.
\cite{Esterling2011,Esterling2011a} rely on hand-coded data by the Congressional Management Foundation, a nonprofit organization which aims to assist Congress. \cite{Therriault2010} use automated text analysis, and also has the most extensive overview of the associated methodology. However, the division of the website into sections (home page, topics, issues, details) is done by hand. The author acquired the websites from the Library of Congress (which only collected them from legislators who actually consented, and Therriault notes that this causes nonrandom missingness).

Important to our methodological objectives, research analyzing and improving the scraping, pre-processing and text analysis pipeline that is applicable to government websites is still in its infancy. \cite{Eschenfelder2002} provide something of an overview of how how federal websites should be assessed from an e-governance point of view, but they largely focus on the substantive criteria that should be fulfilled, rather than the technical aspects of website acquisition and analysis. %In what follows we first define the target dataset---the textual contents of websites of United States municipalities, along with associated metadata on the municipalities and their governments. We then define a pipeline for data collection and analysis that includes methods to access government website URLs, scrape their raw contents from the World Wide Web, gather plain text from the website files, and identify boilerplate text within the plain text contents. Lastly, we illustrate the analysis of municipal government website text by exploring the relationship between the city mayors' party affiliations and the topical contents of the websites.

@@

For data availability reasons, on which we elaborate below, we focus our analysis of municipal websites on six states---Indiana, Louisiana, New York, Washington, California and Texas. These states provide us with a sample that is well-balanced on a number of theoretically important indicators. One, each of the four geographic regions is represented with at least one state. Two, we have a fairly well-balanced sample with respect to the urban/rural cleavage, as both major cities and less densely populated areas are covered. Furthermore, the sample is politically balanced---we have three blue states (CA, WA, NY) and three red states (TX, IN, LA). Finally, our dataset contains some of the wealthiest states (NY, CA, WA and TX are \#2, \#8, \#9 and \#16 respectively, by GDP per capita \citep{BureauofEconomicAnalysis2017}), but also some of the poorer ones (IN and LA).

We acquired the website URLs from two sources: One, we scraped the URLs of city websites from their respective Wikipedia pages, which we found from lists of cities contained within each state. Two, the General Services Administration (GSA) maintains all `.gov' addresses, and provides a complete list of all such domains to the public.\footnote{The dataset is made available at \url{https://github.com/GSA/data/tree/gh-pages/dotgov-domains}. This list is updated once per month---we rely on the version released on January 16, 2017.} The data from the GSA contains the following variables: (1) domain name, specifically, the all-uppercase version of domain and top-level domain (for example, 'ABERDEENMD.GOV'); (2) the type of government entity to which the domain is registered, such as city, county, federal agency, etc; (3) for federal agencies, the name is specified; (4) the city in which the domain is registered. Naturally, the GSA's list does not contain cities which do not use a `.gov' website (or, in many cases, a city owns a registered `.gov' address, but uses a different one). Furthermore, some of the links are non-functional, and some of the county websites on the list are incorrectly marked as city websites (and vice versa). Since the GSA data is less complete and less reliable than the URLs found on Wikipedia, we mainly rely on the former, and only supplement them with the GSA data if a specific city doesn't have a URL recorded on Wikipedia, or our tests (see below) find it to be non-functional.

Not all of the URLS contained in these archives are functional. To test the URLs' functionality, we use a webdriver-controlled browser - a browser that is automatically controlled by a program rather than a human user. We use the Python bindings for the program \texttt{Selenium}, which we use to control \texttt{Firefox} through the webdriver  \texttt{Geckodriver}. This is advantageous compared to conventional scraping tools such as \texttt{Beautiful Soup} or \texttt{Rvest} because most websites are designed to be explored by browsers. Modern browsers perform a lot of actions behind the scenes, such as URL resolution and redirection. The use of a webdriver-controlled browser is necessary in our case because a) some city websites simply don't work, but they don't always output an error code correctly (this can fail, for example, if a webmaster simply stops maintaining a site without removing it entirely) which would throw off an automatic scraper, and more often, b) cities sometimes change their websites' URLs, in which case they redirect from the old to the new URL. A webdriver-controlled browser, unlike the more rigid conventional scraping tools, will simply follow this redirection. This allows us to subsequently record and use the new URL for the actual website scraping. Consequently, an automated browser allows us to robustly answer the following questions: Is the website actually there? Does it work? If not, is it somewhere else or is it broken? We record this information and construct a list of verified URLs.

To download the websites, we rely on the Unix command line tool \texttt{wget}. This program is used to download files from the Internet, and with the use of a recursive option, acts like a webcrawler and scraper. This means that \texttt{wget} downloads HTML files, parses them and then follows the links contained therein. Then it follows those links, and repeats the process until it has constructed a complete tree of the website (note that the program is instructed to stay on the same domain, i.e. it does not follow external links). This way, all the files that make up a website are downloaded. For some cities, whose websites make heavy use of JavaScript to serve content dynamically, such content is not reachable with our methodology and would require additional steps to obtain. For this paper we ignore such sites and restricte our corpus to cities with at least three successfully downloaded pages.

The partisanship of the mayor of each city is coded in different ways, depending on the state. For Indiana, where elections are nominally partisan, this information is accessible through the state government's website\footnote{\url{http://www.in.gov/apps/sos/election/general/general2015?page=office&countyID=1&officeID=32&districtID=-1&candidate=}}. For Louisiana, we received data on the outcomes of mayoral elections from the Local Elections in America Project (LEAP) \citep{marschall2013local}. For the other states, where mayoral elections are not nominally partisan (but the partisanship of the mayor is still well-known), we employed different means: For New York and Washington, we searched the state campaign finance websites, and coded the parties of the candidates based on the party committees from which they received donations. For California and Texas, where our data consists of highly populated cities, partisanship information was acquired from Ballotpedia\footnote{\url{https://ballotpedia.org/List_of_current_mayors_of_the_top_100_cities_in_the_United_States}}. Finally, we also scraped mayoral partisanship from the cities' Wikipedia pages. When compared to the other data sources above, (and manual searches in case of conflicts) Wikipedia proved to be very reliable, and added additional cases to our dataset even for Indiana and Louisiana. Generally speaking, we found data scraped from Wikipedia, aided by manual corrections in case of missing or conflicting data, to be more reliable than data from governmental sources. The partisan breakdown of city websites is depicted in Table \ref{tab:cityparties}. The dataset is has a relative party balance, with variation in each state.

@@

One of the more subtle aspects of local government is the presence of different types of government structures. Between council-manager governments and mayor-council governments \citep{morgan1992policy}---either in the weak or strong mayor variant \citep{desantis2002city}---there is a certain degree of variance in where a city's executive authority lies. Unfortunately we do not have access to information about the type of governments across the breadth of our dataset and therefore cannot explore heterogeneity in the relationship between mayoral partisanship and municipal website contents based on the executive system variant. Given the prominent place that mayors tend to have on their cities' websites, we feel that any bias arising from this nuance should be minor. Information on other covariates (population and median household income - from the American Community Survey 5) was acquired through the API of the U.S. Census Bureau\footnote{\url{https://www.census.gov/data/developers/data-sets.html}}.

@@

Once we have gathered the website files, we have the raw data necessary for text analysis, but it is not yet formatted effectively. As is common in text analysis applications, we need to pre-process our data \citep{denny2018text}. In this section we describe our pre-processing pipeline, with which we take an archive of website files, and output a corpus of formatted plain text files that are suitable for comparative analysis with text as data methods. In this methodological pipeline we address two methodological challenges. First, though they contain significant amounts of text, websites are not comprised of clean plain text files. Rather, the files available at websites are of multiple types, including HTML, PDF, word processor, plain text, and image files. The first step in the methodological pipeline is aimed simply at extracting clean plain text from this heterogeneous file base. The second step in our methodological pipeline is to process the text to remove boilerplate language---language that is effective at differentiating one website from another, but is uninformative regarding policy or politics differences between governments.

\subsection{Site to Text Conversion}
For the most part, the file type of a document can be correctly determined through the filename ending---its extension. However, there are exceptions to this, which, if ignored, can lead to large amounts of improperly formatted text, arising from incorrectly converted documents, which leads to a general decrease in the amount of usable data. Two issues in particular need to be addressed: One, HTML files on city websites frequently do not have an ending, but are still perfectly readable if correctly identified as such. Second, some documents contain the incorrect file ending. For example, we found thousands of documents that ended in .html, when they were actually PDFs. To accurately assess their type, we rely on the R package \texttt{wand} \citep{wand}, which is an R interface to the Unix library \texttt{libmagic} \citep{darwin2008libmagic}, which determines the type of a file on the basis of its file signature - or ``magic number''. This short sequence of bytes at the start (and sometimes end) of files is unique for each file type and therefore allows its correct identification through computer forensics tools such as \texttt{libmagic}.

Consequently we rename all documents so that their file ending reflects their actual file type. This is strictly necessary, because we rely on the \texttt{readtext} R package \citep{readtext}, which determines a document's type solely through its ending---to convert the files to plain text.\footnote{We have also experimented with several Unix-based alternatives, but found that they largely led to the same results as \texttt{readtext}.} The breakdown of the files by type is given in Table \ref{tab:filetypes}. The most frequent file types are HTML and PDF, from which we are able to extract a substantial amount of usable text. Files of type XML, DOC, TXT, and DOCX, also occur regularly in our corpus and offer a considerable volume of textual data.

@@

We then take several steps to pre-processing the data as required for subsequent the subsequent analysis. Pre-processing choices should be contingent on the analysis being conducted with the text later and can have significant effects on the outcomes of an analysis \citep{denny2018text}. The steps taken here should therefore not be seen as universally applicable to all analyses of government website data.

The text documents are converted to UTF-8 and then stripped of dates, punctuation, numbers and words connected by underscores. At this point, the documents of one city still closely resemble one another in the form of boilerplate content, be it website elements (i.e. "You are here", "Home", "Directory" etc.) in html documents, or commonly used forms or phrases in pdfs, doc and docx files. This is an issue, because this boilerplate content causes the results of analyzing this data with text analysis methods to characterize documents primarily by the cities from which they originate (through their unique boilerplate structure, e.g. a menu with certain terms repeated on every site of the domain), and not the substantive features of their contents. Our solution to this problem is described in more detail in Section \ref{boilerplate}.

>>>>

Local governments convey voluminous information about all aspects of their policymaking, policy implementation, and public deliberation, via their official websites. The vital role of official websites in connecting the government and the governed has motivated a wave of research on the contents of government websites \citep[e.g.,][]{grimmelikhuijsen2010transparency,wang2005evaluating,osman2014cobra}. Despite the potential for automated scraping of website contents, the conventional approach to data collection in projects focused on government websites involves manual content extraction from each website in the dataset. Though highly accurate, the manual approach to data collection is costly, and cannot be scaled to capture even a fraction of the volume of content available on government websites. In this paper, we present a methodological pipeline that can be used to automatically scrape government websites in order to build datasets that can be used for text analysis. We provide an illustrative application in which we explore the ways in which the textual contents on city government websites in six American states (IN, LA, NY, WA, CA and TX) correlate with the partisanship of the city mayor.

Though there exists a variety of software tools that are designed to automatically scrape all of the files available at a website \citep{glez2013web}, raw website downloads have to be processed significantly before the files are adequately prepared for content (e.g., text, image) analysis. We describe and provide solutions to two central challenges in automatically gathering and analyzing website textual contents. First, plain text must be extracted from the files. This involves purging the files of syntax in HTML or other markup and programming languages and discarding any character encoding errors that result from reading in files with incorrectly specified character encodings. This challenge would arise in any context in which researchers sought to study the textual contents of websites, and is not unique to comparative analysis of government websites. The second challenge we address in our methodological pipeline is, however, specific to the research objective of comparing websites on the basis of a common lexicon. For any two governments, the textual signatures that most dramatically differentiate the textual contents of their websites consist of what we can call ``boilerplate'' text---header, footer, or other titling text that is designed to identify the website as being associated with a specific government entity (e.g., ``Welcome to the city of Santa Cruz'', ``The City of Los Angeles welcomes you''). This boilerplate text is replicated across many files that are associated with a government's website, but it provides little information regarding the form and/or function of the government. The second methodological innovation we offer in our pipeline is designed to minimize the impact of this boilerplate text on the comparative analysis of government website content.

Government websites provide information about how public policies shape the lives of local residents, and how local residents can engage with government to shape public policy. As such, government websites reflect both the results of and inputs to, the political leadership in the city. In our illustrative application, we explore the ways in which the contents of city government websites differ on the basis of the partisanship of the city's elected executive. A substantial body of research has found that the partisanship of the mayor affects city governance along multiple dimensions, including city budget priorities \citep{de2016mayoral}, policies affecting inequality in cities \citep{einstein2016mayors}, and framing of criminal justice policy \citep{marion2013mayor}. Furthermore, recent media coverage of changes to government websites that follow transitions in party control suggest that changes in web content are salient government actions, as perceived by the general public \citep{sharfstein2017science,kirby2017trump,duarte2017deniable}. We study whether significant differences between city governments based on mayoral partisanship are reflected in the contents of city websites.

@@

 According to \cite{Mayhew1974}, politicians engage in advertising, credit claiming and position taking in order to get re-elected. Official city websites allow mayors to perform all three of these functions. Their offices frequently take a prominent position on the front page, and many websites also feature a picture of the mayor. We present an example of this in Figure \ref{fig:eriemayor}. The Erie, Pennsylvania website homepage presents an image of the Democratic mayor, Joseph Schember, along with a list of laudable attributes of the city. In local politics, where campaign funds are low, this lends incumbents a crucial advantage in becoming more well-known among their constituencies. Furthermore, municipal politics gives incumbents clear and tangible achievements they can point to, such as completed infrastructure projects, the acquisition of federal or state funding, or the hosting of city-wide events. City websites present an opportunity for local officials to brandish these accomplishments. Finally, they also give mayors a platform from which they can advertise their political beliefs. On municipal websites, this may not manifest in the form of brazen partisanship, but more subtle avenues are available. As noted by \cite{einstein2016mayors}, there are stark differences in the spending preferences of Democratic and Republican mayors. City websites can then be used to communicate the stance of a mayor on social or economic programs. Another advantage of websites with regard to communication is that unlike direct social interactions, officials have full control over them.

@@

The existing research that uses scraped websites provides an indication of the theoretical value of empirical analysis of web contents produced by governments, public officials, and candidates for office. The most pertinent literature to our research is the e-governance literature, which focuses on the online presence of governments from a usability and public service point of view. For the most part, research in this category develops a classification scheme to rate websites in terms of accessibility, ease-of-use, and function, and then hand-codes a set of websites according to these criteria \citep[e.g., ][]{Urban2002,Armstrong2011,Feeney2017}. As an example, \cite{grimmelikhuijsen2012developing} study local government websites with the goal of uncovering how they aid the goal of transparency. To this end, they analyze a set of Dutch municipalities in which air quality had deteriorated. The authors test whether local governments provide citizens with information about potential complications and solutions associated with this issue. Like most e-government studies, however, this publication does not make any use of automated text analysis.
%Websites have also played a major role in the field of media studies, as scholars have scraped and analyzed the online presence of newspapers, as well as the more diffuse world of online political blogs \citep{Adamic2005,Gentzkow2010}. \cite{Lin2011} provide a good example of a study which makes extensive use of automated content analysis - a necessity arising from its dataset of 66830 blog posts and 57221 online news articles. The authors estimate the political slant of these entities by counting the frequencies with which politicians of either side are mentioned and determine that blogs are generally more biased. Unfortunately for us, the authors don't go into the details of their text analysis, and offer no information on the acquisition and pre-processing of the data.
%Another well-known example fitting into this area of study is the set of studies conducted by King et al. \citep{KING2013,King2014,KING2017}, in which the authors study censorship by the country's government on its lively blogosphere. However, the authors also provide no information on how their data was collected ``our extensive engineering effort, which we do not detail here for obvious reasons [...]''.

The websites of politicians and their parties have also fallen under scholarly scrutiny. Researchers have found that in order to identify the constituencies, motives, and modes of communication among these actors, their websites can be very illuminating sources of information \citep{Druckman2009,Druckman2010,Cryer2017,Esterling2011,Esterling2011a,Norris2003,Therriault2010}. \cite{Druckman2009,Druckman2010} rely on the \textit{National Journal} to find the websites, then hand-coded them. \cite{Cryer2017} mentions that she relied on Archive-it, a web service of the Internet Archive. Though the Internet Archive may provide extensive coverage of high profile officials or national governments, we found that its coverage of municipal government websites was sparse and irregular.
\cite{Esterling2011,Esterling2011a} rely on hand-coded data by the Congressional Management Foundation, a nonprofit organization which aims to assist Congress. \cite{Therriault2010} use automated text analysis, and also has the most extensive overview of the associated methodology. However, the division of the website into sections (home page, topics, issues, details) is done by hand. The author acquired the websites from the Library of Congress (which only collected them from legislators who actually consented, and Therriault notes that this causes nonrandom missingness).

Important to our methodological objectives, research analyzing and improving the scraping, pre-processing and text analysis pipeline that is applicable to government websites is still in its infancy. \cite{Eschenfelder2002} provide something of an overview of how federal websites should be assessed from an e-governance point of view, but they largely focus on the substantive criteria that should be fulfilled, rather than the technical aspects of website acquisition and analysis. %In what follows we first define the target dataset---the textual contents of websites of United States municipalities, along with associated metadata on the municipalities and their governments. We then define a pipeline for data collection and analysis that includes methods to access government website URLs, scrape their raw contents from the World Wide Web, gather plain text from the website files, and identify boilerplate text within the plain text contents. Lastly, we illustrate the analysis of municipal government website text by exploring the relationship between the city mayors' party affiliations and the topical contents of the websites.

@@

For data availability reasons, on which we elaborate below, we focus our analysis of municipal websites on six states---Indiana, Louisiana, New York, Washington, California, and Texas. These states provide us with a sample that is well-balanced on a number of theoretically important indicators. One, each of the four geographic regions are represented with at least one state. Two, we have a fairly well-balanced sample with respect to the urban/rural cleavage, as both major cities and less densely populated areas are covered. Furthermore, the sample is politically balanced---we have three blue states (CA, WA, NY) and three red states (TX, IN, LA). Finally, our dataset contains some of the wealthiest states (NY, CA, WA and TX are \#2, \#8, \#9 and \#16 respectively, by GDP per capita \citep{BureauofEconomicAnalysis2017}), but also some of the poorer ones (IN and LA).

We acquired the website URLs from two sources: One, we scraped the URLs of city websites from their respective Wikipedia pages, which we found from lists of cities contained within each state. Two, the General Services Administration (GSA) maintains all `.gov' addresses, and provides a complete list of all such domains to the public.\footnote{The dataset is made available at \url{https://github.com/GSA/data/tree/gh-pages/dotgov-domains}. This list is updated once per month---we rely on the version released on January 16, 2017.} The data from the GSA contains the following variables: (1) domain name, specifically, the all-uppercase version of domain and top-level domain (for example, 'ABERDEENMD.GOV'); (2) the type of government entity to which the domain is registered, such as city, county, federal agency, etc; (3) for federal agencies, the name is specified; (4) the city in which the domain is registered. Naturally, the GSA's list does not contain cities which do not use a `.gov' website (or, in many cases, a city owns a registered `.gov' address, but uses a different one). Furthermore, some of the links are non-functional, and some of the county websites on the list are incorrectly marked as city websites (and vice versa). Since the GSA data is less complete and less reliable than the URLs found on Wikipedia, we mainly rely on the former and only supplement them with the GSA data if a specific city doesn't have a URL recorded on Wikipedia, or our tests (see below) find it to be non-functional.

Not all of the URLs contained in these archives are functional. To test the URLs' functionality, we use a web driver-controlled browser - a browser that is automatically controlled by a program rather than a human user. We use the Python bindings for the program \texttt{Selenium}, which we use to control \texttt{Firefox} through the web driver  \texttt{Geckodriver}. This is advantageous compared to conventional scraping tools such as \texttt{Beautiful Soup} or \texttt{Rvest} because most websites are designed to be explored by browsers. Modern browsers perform a lot of actions behind the scenes, such as URL resolution and redirection. The use of a web drive-controlled browser is necessary in our case because a) some city websites simply don't work, but they don't always output an error code correctly (this can fail, for example, if a webmaster simply stops maintaining a site without removing it entirely) which would throw off an automatic scraper, and more often, b) cities sometimes change their websites' URLs, in which case they redirect from the old to the new URL. A web driver-controlled browser, unlike the more rigid conventional scraping tools, will simply follow this redirection. This allows us to subsequently record and use the new URL for the actual website scraping. Consequently, an automated browser allows us to robustly answer the following questions: Is the website actually there? Does it work? If not, is it somewhere else or is it broken? We record this information and construct a list of verified URLs.

To download the websites, we rely on the Unix command line tool \texttt{wget}. This program is used to download files from the Internet, and with the use of a recursive option, acts like a web crawler and scraper. This means that \texttt{wget} downloads HTML files, parses them and then follows the links contained therein. Then it follows those links and repeats the process until it has constructed a complete tree of the website (note that the program is instructed to stay on the same domain, i.e. it does not follow external links). This way, all the files that make up a website are downloaded. For some cities, whose websites make heavy use of JavaScript to serve content dynamically, such content is not reachable with our methodology and would require additional steps to obtain. For this paper, we ignore such sites and restricted our corpus to cities with at least three successfully downloaded pages.

The partisanship of the mayor of each city is coded in different ways, depending on the state. For Indiana, where elections are nominally partisan, this information is accessible through the state government's website\footnote{\url{http://www.in.gov/apps/sos/election/general/general2015?page=office&countyID=1&officeID=32&districtID=-1&candidate=}}. For Louisiana, we received data on the outcomes of mayoral elections from the Local Elections in America Project (LEAP) \citep{marschall2013local}. For the other states, where mayoral elections are not nominally partisan (but the partisanship of the mayor is still well-known), we employed different means: For New York and Washington, we searched the state campaign finance websites, and coded the parties of the candidates based on the party committees from which they received donations. For California and Texas, where our data consists of highly populated cities, partisanship information was acquired from Ballotpedia\footnote{\url{https://ballotpedia.org/List_of_current_mayors_of_the_top_100_cities_in_the_United_States}}. Finally, we also scraped mayoral partisanship from the cities' Wikipedia pages. When compared to the other data sources above, (and manual searches in case of conflicts) Wikipedia proved to be very reliable and added additional cases to our dataset even for Indiana and Louisiana. Generally speaking, we found data scraped from Wikipedia, aided by manual corrections in case of missing or conflicting data, to be more reliable than data from governmental sources. The partisan breakdown of city websites is depicted in Table \ref{tab:cityparties}. The dataset is has a relative party balance, with variation in each state.

@@

One of the more subtle aspects of local government is the presence of different types of government structures. Between council-manager governments and mayor-council governments \citep{morgan1992policy}---either in the weak or strong mayor variant \citep{desantis2002city}---there is a certain degree of variance in where a city's executive authority lies. Unfortunately, we do not have access to information about the type of governments across the breadth of our dataset and therefore cannot explore heterogeneity in the relationship between mayoral partisanship and municipal website contents based on the executive system variant. Given the prominent place that mayors tend to have on their cities' websites, we feel that any bias arising from this nuance should be minor. Information on other covariates (population and median household income - from the American Community Survey 5) was acquired through the API of the U.S. Census Bureau\footnote{\url{https://www.census.gov/data/developers/data-sets.html}}.

@@

Once we have gathered the website files, we have the raw data necessary for text analysis, but it is not yet formatted effectively. As is common in text analysis applications, we need to pre-process our data \citep{denny2018text}. In this section, we describe our pre-processing pipeline, with which we take an archive of website files, and output a corpus of formatted plain text files that are suitable for comparative analysis with text as data methods. In this methodological pipeline, we address two methodological challenges. First, though they contain significant amounts of text, websites are not comprised of clean plain text files. Rather, the files available at websites are of multiple types, including HTML, PDF, word processor, plain text, and image files. The first step in the methodological pipeline is aimed simply at extracting clean plain text from this heterogeneous file base. The second step in our methodological pipeline is to process the text to remove boilerplate language---language that is effective at differentiating one website from another but is uninformative regarding policy or political differences between governments.

\subsection{Site to Text Conversion}
For the most part, the file type of a document can be correctly determined through the filename ending---its extension. However, there are exceptions to this, which, if ignored, can lead to large amounts of improperly formatted text, arising from incorrectly converted documents, which leads to a general decrease in the amount of usable data. Two issues, in particular, need to be addressed: One, HTML files on city websites frequently do not have an ending but are still perfectly readable if correctly identified as such. Second, some documents contain the incorrect file ending. For example, we found thousands of documents that ended in .html, when they were actually PDFs. To accurately assess their type, we rely on the R package \texttt{wand} \citep{wand}, which is an R interface to the Unix library \texttt{libmagic} \citep{darwin2008libmagic}, which determines the type of a file on the basis of its file signature - or ``magic number''. This short sequence of bytes at the start (and sometimes end) of files is unique for each file type and therefore allows its correct identification through computer forensics tools such as \texttt{libmagic}.

Consequently, we rename all documents so that their file ending reflects their actual file type. This is strictly necessary because we rely on the \texttt{readtext} R package \citep{readtext}, which determines a document's type solely through its ending---to convert the files to plain text.\footnote{We have also experimented with several Unix-based alternatives, but found that they largely led to the same results as \texttt{readtext}.} The breakdown of the files by type is given in Table \ref{tab:filetypes}. The most frequent file types are HTML and PDF, from which we are able to extract a substantial amount of usable text. Files of type XML, DOC, TXT, and DOCX, also occur regularly in our corpus and offer a considerable volume of textual data.

@@

We then take several steps to pre-process the data as required for subsequent the subsequent analysis. Pre-processing choices should be contingent on the analysis being conducted with the text later and can have significant effects on the outcomes of an analysis \citep{denny2018text}. The steps taken here should therefore not be seen as universally applicable to all analyses of government website data.

The text documents are converted to UTF-8 and then stripped of dates, punctuation, numbers, and words connected by underscores. At this point, the documents of one city still closely resemble one another in the form of boilerplate content, be it website elements (i.e. "You are here", "Home", "Directory" etc.) in HTML documents, or commonly used forms or phrases in pdf, doc and docx files. This is an issue, because this boilerplate content causes the results of analyzing this data with text analysis methods to characterize documents primarily by the cities from which they originate (through their unique boilerplate structure, e.g. a menu with certain terms repeated on every site of the domain), and not the substantive features of their contents. Our solution to this problem is described in more detail in Section \ref{boilerplate}.

Preprocessing further includes setting every character to lowercase, as well as the removal of bullet points which frequently occur in HTML documents, extraneous whitespace, XML documents mislabeled as HTML files, and empty documents. Furthermore, some documents contain gibberish, often as a result of faulty or impartial optical character recognition applied to text that was produced through a non-machine-readable medium. To combat this problem, we employ two solutions. One, we use spellchecking, implemented through the \texttt{hunspell} R package \citep{hunspell}, to remove all non-English words.\footnote{Some of the cities, for example, Los Angeles, do contain a sizable proportion of Spanish content. The analysis of this content is beyond the scope of this paper but could be explored in future work, for example using methods of text processing that are applicable to multilingual corpora \citep{lucas2015computer}. } However, \texttt{hunspell} does not cover everything, either because some tokens are not actual words (for example artifacts from defective encoding), or because random sequences of characters just so happen to form words that exist in a dictionary (for example "eh" or "duh"). Since we rely on a bag-of-words model in which syntax does not matter, we can ameliorate these problems by removing all text except for whitespaces and the characters that appear in the English alphabet. Since a lot of the nonsensical text tends to be quite repetitive, we also delete all documents in which the proportion of unique to the total number of tokens is less than 0.15. Furthermore, \texttt{hunspell} does not spellcheck individual characters or two-character words, so we remove these token types entirely. Since these pre-processing steps reduce documents which are largely unsuitable to only a few tokens (i.e., word occurrences), we also remove all remaining documents containing less than 50 tokens. Finally, to remove words that are extremely rare (which also has the advantage of eliminating any remaining oddities) and thus add nothing substantive to our models while increasing their computational cost, we also discard any token types that occur in only one document. We also conduct lemmatization to reduce words to their basic form. Lemmatization is similar to stemming but works in a somewhat more sophisticated manner by taking grammar and surrounding words into account to identify the dictionary form of a word. For example, the lemma of the word ``lemmatization'' would be ``lemmatize'', whereas most stemmers would simply chop off the ending, which would yield ``lemmatiz''. Thus, lemmatization makes the results more easily comprehensible. To this end, we rely on the R package \texttt{spacyr}, which provides an R implementation of the Python library \texttt{spaCy}.
